---
title: "ds202_lab5"
author: "Sam Begley"
date: "April 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv")
pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv")
library(tidyverse)
```
##Part 1)


```{r 1}
head(acc)
names(acc)
acc$DAY_WEEK <- as.factor(acc$DAY_WEEK)
levels(acc$DAY_WEEK)
levels(acc$DAY_WEEK) <- c('Sun', 'Mon', 'Tues', 'Wed', 'Thur', 'Fri', 'Sat')
acc %>% group_by(DAY_WEEK) %>% 
  summarise(count = n()) %>%
  ggplot(aes(x = DAY_WEEK, y = count)) +
  geom_bar(stat = 'identity')
acc %>% group_by(HOUR) %>%
  filter(HOUR != 99) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = HOUR, y = count)) +
  geom_bar(stat = 'identity')
acc %>% filter(DRUNK_DR != 0) %>%
  summarise(count = n())
```


Accidents occur on Saturdays more often than any other day of the week. There seems to be more accidents on weekends than on weekdays.

6:00 PM is the most common time for accidents to occur, which makes sense because that is around rush hour, when people are leaving work. 4:00 AM is the least common time for accidents to occur, which also makes since there aren't usually many cars on the road at that time.

There were 8,769 accidents with at least 1 drunk driver.

##Part 2)


```{R 2}
drivers <- pers %>% filter(PER_TYP == 1)
intersect(names(acc), names(drivers))
drivers <- drivers %>% select('ST_CASE', 'SEX')
data <- left_join(acc, drivers, by = 'ST_CASE')
head(data)
data$DAY_WEEK <- as.factor(data$DAY_WEEK)
levels(data$DAY_WEEK) <- c('Sun', 'Mon', 'Tues', 'Wed', 'Thur', 'Fri', 'Sat')
data$SEX <- as.factor(data$SEX)
levels(data$SEX) <- c('Male', 'Female', 8, 9)
data %>% filter(HOUR != 99) %>%
  group_by(DAY_WEEK, HOUR, SEX) %>%
  summarise(COUNT = n()) %>%
  filter(SEX != 8 & SEX != 9) %>%
  ggplot(aes(x = HOUR, y = COUNT, color = SEX)) +
  geom_line() +
  facet_grid(DAY_WEEK ~ .)
```


##Part 3)


```{r 3}
counties <- map_data('county')
names <- readxl::read_xlsx('~/SD_Shortcut/FRPP GLC United States.xlsx')
names(names) <- c('Territory', 'region', 'State Code', 'City Code', 'City Name', 'COUNTY', 'subregion', 'Country Code', 'Old City Name', 'Date Redord Added')
names <- names[-1, ]
names <- names %>% select('region', 'COUNTY', 'subregion')
counties$region <- toupper(counties$region)
counties$subregion <- toupper(counties$subregion)
names <- names %>% mutate(tregion = paste0(region, subregion))
names <- names %>% distinct(tregion, .keep_all = TRUE)
counties <- counties %>% mutate(tregion = paste0(region, subregion))
data2 <- left_join(counties, names, by = 'tregion')
data2$COUNTY <- as.integer(data2$COUNTY)
data3 <- data %>% group_by(COUNTY) %>%
  summarise(count = n_distinct(ST_CASE))
data4 <- left_join(data2, data3, by = 'COUNTY')
data4 <- data4 %>% distinct(long, lat, group, 
                            .keep_all = TRUE)

data4 %>% ggplot(aes(x = long, y = lat)) +
  geom_polygon(aes(group = group, fill = count))
counties %>% ggplot(aes(x = long, y = lat)) +
  geom_polygon(aes(group = group))
```
